Quality of Service with Software Defined Network.

### 2017-03-07 report/paper.tex
altre aggiunte

\section{Introduction}\label{sec:intro}
Software-defined networking (SDN) is way to manage networks that separates the control plane from the forwarding plane.
SDN offers a centralized view of the network, giving an SDN Controller the ability to act as the “brains” of the network.
The SDN Controller relays information to switches and routers via southbound APIs, and to the applications with northbound APIs.
One of the most well-known protocols used by SDN Controllers is OpenFlow, however, it isn’t the only SDN standard, but it is widely accepted.
Centralized, programmable SDN environments can easily adjust to the rapidly changing needs of businesses. SDN can lower costs and limit
wasteful provisioning, as well as provide flexibility and innovation for networks.
\#TODO da completare
  \subsection{OpenFlow}
  During 2006 thanks to M. Casado was developed Ethane a new network architecture on this conncept during 2011 was bord the Open Network Foundatin.
  This is a no profit association that would innovate SDN and standardize OpenFlow protocol whit relatives tecnology \cite{ONF}.
  The first version of OpenFlow is 1.0.0 release on 2009 but the last standardized version is v1.5.1 approved on march 2015 \cite{ONF_report}.
  \#TODO da sistemare
  \\
  \\
  \textbf{Openflow 1.0 messages}
Openflow defines a number of messages between a controller and a switch. Now there will be presented some messages we exploited:\cite{openflow}
\begin{itemize}

@@

@@

\section{Instrument}\label{sec:instrument}
\#TODO matteo
  \subsection{Controller}
  The controller that we use is POX, the software is opensource and it is write with Python, it is based on SDN and support OpenFlow.
  The software use the version 2.7 of Python.
  It's repository is at this link \url{https://github.com/noxrepo/pox} and a lot of documentation about the progect
  is present at this link \url{https://openflow.stanford.edu/display/ONL/POX+Wiki}.
  The only problem to use this controller is that it support only the first version of OpenFlow1.0.
  \\
  The controller run inside a VirtualMachine on our personal computer,
  the VM is offered by SDN hub on this site: \url{http://sdnhub.org/tutorials/sdn-tutorial-vm/}.
  This VM is a 64-bit Ubuntu 14.04 that has a number of SDN software and tools installed.
  % \#TODO matteo da correggere e completare
  \subsection{Router Board}
  \begin{figure}[!h]
 \centering
 \includegraphics[scale=0.30]{images/rboard.png}
 \caption{\emph{Experience scenario}}
 \label{fig:topo}
  \end{figure}

@@

\section{Results}\label{sec:results}
\subsection{Real Lab Scenario}
The topological scenario in which we have set up experience is the next:
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/topo.png}
 \caption{\emph{Experience scenario}}
 \label{fig:topo}
\end{figure}

We have used some principal function released with POX controller:
\begin{itemize}
 %\item \texttt{pox.topology.launch()} and \texttt{pox.openflow.topology.launch()} descrivere a che ci serve\cite{pox}.
 \item \texttt{pox.openflow.discovery.launch()} this function run algorithm in order to discovery the network and every time in witch some new 
 topological change occour it return an event with ( con cui ) we can update our topological knowledge.
 It does not provide the host event. For host event\cite{pox}.
 \item \texttt{pox.openflow.spanning_tree.launch()} this function is necessary with a presence of loop on topology 
 because with a spanning_tree algorithm it avoid the broadcast storm of message. When we set up \texttt{OFPP_FLOOD} like output port action in a table's 
 switch row if some spanning_tree algorithm run the \texttt{OFPP_FLOOD} is only theoretical infact the algorithm close some port for flooding service 
 on the switch in orther to avoid loop\cite{pox}.
\end{itemize}

\#TODO %qui non so se serve un'introduzione o meno
  \subsection{Graph and Network structures}
  In order to achieve the realtime adaptability and statistics storage we have develop ad-hoc structure.
  We use a specific python library called NetworkX \cite{networkx} to create a graph in witch we can storage ad-hoc node that represent
  resplectively switch or host element.
  This topological structure is updated realtime thanks to connection event and link discover offered by various library.
  ( mettere i vari eventi che sfrutto )
  The link besides have many parameters in witch we can save network performance. To update every link parameter we have to creat ad hoc function
  because there isn't some function constructed yet. ( sistemare frase) .
  This type of structure allow us to run Dijkstra algorithm that return the minimum path ( in base a vari parametri ) on a nlog n time ( verificare).
  (aggiungere qualcosa e correggere)
  \#TODO matteo

  \subsection{Traffics analisis}\label{subsec:traf}
  For every networks performance we have tryed to develop a specific solution. The OpenFlow main goal is to centralize controll and software
  complexity so the routerboard can made only some easy function in orther to contain the hardware complexity and mantainly a good speed\cite{qos_paper}.
  Il che ha portato a dover utilizzare stratagemmi.
  \#TODO %riformulare
    \subsubsection{Delay}
Monitoring round-trip time provides important insights for network troubleshooting and traffic engineering.
The  common  monitoring  technique  is  to  actively  send  probe packets  from  selected  vantage  points  (hosts  or  middleboxes).
In traditional  networks, the control over the network routing is
limited,  making  it  impossible  to  monitor  every  selected  path. The  emerging  concept  of  Software  Defined  Networking  sim-plifies 
network  control.  However,  OpenFlow,  the  common  SDN protocol,   does   not   support   RTT   monitoring   as   part   of   its
specification. In this experience we have tryed to emply an tecnic to escape this problem\cite{rtt}.
\newline The idea is to measure the time that occure from sending a packet from controller to a switch and the time in which the packet return from 
an other switch.
\\
\newline First we have to discovery round trip time from controller to every single switch on our topology.
In order to do this on the controller we start a timer, we send a proper packet to the switch specifing the \texttt{OFPP_CONTROLLER - Send to the controller}
option and when the message return to the controller we stop the timer and calculare RTT for the switch.
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/rtt2.png}
 \caption{\emph{RTT between controller and OF switch}}
 \label{fig:topo}
\end{figure}
\newline
If we need to calculate the delay on a specific link, the rtt has to calculate with the headed switch of this link.
\\
\newline 
Now we create a proper message with unique identifier number, we start a timer \texttt{TM1} and then send this message to a first switch specifing the output port.
Then when the packet comes to the second switch it identify the particular message and sends to the controller that stops it's dedicated timer \texttt{TM2}.
\newline This part required a particolar flow entry on the second switch in orther to send to the controller only some particulare packet.
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/rtt1.png}
 \caption{\emph{Pck from controller to itself following a particular way.}}
 \label{fig:topo}
\end{figure}
\\
\newline At this point we can calculate the delay of the link.
\\
\\
 delay = (TM2 - TM1) - RTT1/2 - RTT2/2
\\
\\
Teorethically this method must works but in practice there are a lot of delay introduce from computational concestion on the virtualized controller.
What we observe is a estimated delay like some seconds that is impossible, infact if we made some \texttt{ping test} from one host to another host 
in this SDN network we realize delay around 3ms.
\\

>>>>

\section{Introduction}\label{sec:intro}
Software-defined networking (SDN) is way to manage networks that separates the control plane from the forwarding plane.
SDN offers a centralized view of the network, giving an SDN Controller the ability to act as the “brains” of the network.
The SDN Controller relays information to switches and routers via southbound APIs, and to the applications with northbound APIs.
One of the most well-known protocols used by SDN Controllers is OpenFlow, however, it isn’t the only SDN standard, but it is widely accepted.
Centralized, programmable SDN environments can easily adjust to the rapidly changing needs of businesses. SDN can lower costs and limit
wasteful provisioning, as well as provide flexibility and innovation for networks.
\#TODO da completare
  \subsection{OpenFlow}
  During 2006 thanks to M. Casado was developed Ethane a new network architecture on this conncept during 2011 was bord the Open Network Foundatin.
  This is a no profit association that would innovate SDN and standardize OpenFlow protocol whit relatives tecnology \cite{ONF}.
  The first version of OpenFlow is 1.0.0 release on 2009 but the last standardized version is v1.5.1 approved on march 2015 \cite{ONF_report}.
  \#TODO da sistemare
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.58]{images/of.jpg}
 \caption{\emph{OpenFlow}}
 \label{fig:topo}
\end{figure}
  \\
  \textbf{Openflow 1.0 messages}
Openflow defines a number of messages between a controller and a switch. Now there will be presented some messages we exploited:\cite{openflow}
\begin{itemize}

@@

\subsection{Known Limitations}
SDN networks are known to be underperforming with low traffic, which means that if nodes tries to connect with each other
to exchange little traffic, the network will have a higher delay and will perform worst than a
traditional network. We tried to reduce this underperformance with the usage of the wildcards:
Every rule needs at least the source ip or the destination ip. If one of them is in a network
(e.g. the internet traffic is in the network 0.0.0.0/0 excluded 192.168.10.0/24 ),
we can redirect all the traffic from an internal host to the internet with a predefined
path. This means that only one PacketIn will be sent to the controller if a host is surfing the network.

Another limitation is the no default routing from a source to a destination: when an host
is plugged to a network of SDN switches, it won't know the destinatio. The only way it can
discover the host is by flooding the network with that packet, but this makes each switch raise a PacketIn message to the controller
slowing down the startup of the connection. This thing doesn't happen with traditional routing, because each router knows
the network disposition, but doesn't know each host how much traffic is doing. For that the administrator need to put a firewall.
The last one isn't needed, as the controller can set a switch to drop all the packets from/to a destination.

@@

\section{Instrument}\label{sec:instrument}
  \subsection{Controller}
  The controller that we use is POX, the software is opensource and it is write with Python, it is based on SDN and support OpenFlow.
  The software use the version 2.7 of Python.
  It's repository is at this link \url{https://github.com/noxrepo/pox} and a lot of documentation about the progect
  is present at this link \url{https://openflow.stanford.edu/display/ONL/POX+Wiki}.
  The only problem to use this controller is that it support only the first version of OpenFlow1.0.
  \\
  The controller run inside a VirtualMachine on our personal computer,
  the VM is offered by SDN hub on this site: \url{http://sdnhub.org/tutorials/sdn-tutorial-vm/}.
  This VM is a 64-bit Ubuntu 14.04 that has a number of SDN software and tools installed.
  % \#TODO matteo da correggere e completare
  \subsection{Router Board}
  \begin{figure}[!h]
 \centering
 \includegraphics[scale=0.30]{images/rboard.png}
 \caption{\emph{Experience scenario}}
 \label{fig:topo}
  \end{figure}

@@

\section{Results}\label{sec:results}
\subsection{Controller:}
We developed a controller that can handle a mesh network, which has loops, that avoid
the traffic flooding. The controller stores the switch and the connection between them;
knows which hosts are connected to each switch port, with the hypothesys that only one
host is connected at each port, unless the port is the one connected to the internet.
With the use of the statistics, the controller get the status of the network and, if necessary,
tries to adjust the traffic and put in the link with the higher path loss the torrent traffic,
make the connection between two gaming host the one with less delay and tries to redistribute all the traffic in a way
such that all the nodes have an average throughput as lowest as possible.
\\
\newline
We have used some principal function released with POX controller:
\begin{itemize}
 %\item \texttt{pox.topology.launch()} and \texttt{pox.openflow.topology.launch()} descrivere a che ci serve\cite{pox}.
 \item \texttt{pox.openflow.discovery.launch()} this function run algorithm in order to discovery the network and every time in witch some new 
 topological change occour it return an event with ( con cui ) we can update our topological knowledge.
 It does not provide the host event. For host event\cite{pox}.
 \item \texttt{pox.openflow.spanning_tree.launch()} this function is necessary with a presence of loop on topology 
 because with a spanning_tree algorithm it avoid the broadcast storm of message. When we set up \texttt{OFPP_FLOOD} like output port action in a table's 
 switch row if some spanning_tree algorithm run the \texttt{OFPP_FLOOD} is only theoretical infact the algorithm close some port for flooding service 
 on the switch in orther to avoid loop\cite{pox}.
\end{itemize}
\subsection{Real Lab Scenario}
The topological scenario in which we have set up experience is the next:
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/topo.png}
 \caption{\emph{Experience scenario}}
 \label{fig:topo}
\end{figure}

We have use 4 MikroTik Router board that were provided us in the Communication Networks Lab at DEI in University of Padua.
The 4 routerboard is connected each other in order to design a loop. Every sigle board connect two host and is connected with a legacy switch that provide 
to logically connected with controller.

\subsection{Graph and Network structures}
  In order to achieve the realtime adaptability and statistics storage we have develop ad-hoc structure.
  We use a specific python library called NetworkX \cite{networkx} to create a graph in witch we can storage ad-hoc node that represent
  resplectively switch or host element.
  This topological structure is updated realtime thanks to connection event and link discover offered by various library.
  ( mettere i vari eventi che sfrutto )
  The link besides have many parameters in witch we can save network performance. To update every link parameter we have to creat ad hoc function
  because there isn't some function constructed yet. ( sistemare frase) .
  This type of structure allow us to run Dijkstra algorithm that return the minimum path ( in base a vari parametri ) on a nlog n time ( verificare).
  (aggiungere qualcosa e correggere)
  \#TODO matteo

  \subsection{Traffics analisis}\label{subsec:traf}
  We used laptop connected via
GigabitEthernet interface and Ethernet cable cat.5a to receive the traffic from the destination host.
Our laptop has been connected via Wi-Fi to our access point via protocol 802.11n.
No other traffic sources were active during the testing, so the router was handling only
the traffic from laptop when sniffing traffic. The connection used was an ADSL2+ via copper.
We did 3 type of traffic analysis: Gaming traffic, Skype® and torrent traffic. We analyzed
\#Mb for the gaming traffic, \#Mb for the VoIP traffic and 3GB for the torrent traffic.
The analysis involved the average number of packets per second, the packet size, the protocol (TCP or UDP)

The traffic analysis of the torrent protocol shows that the average packet size is the MTU
(i.e. 1500bytes) and the protocol tryies to exploit all the channel capabilities. The traffic
is made via UDP packets with the port range in the ephimeral ports (2000_- 65535).
Moreover, the node is linked to a lot of IPs with multiple connection: this means that
if someone wants to find a p2p traffic, it can look to the statistics and see if the traffic is loading all the channel,
to how many IPs, an host is connected to and if it uses UDP.
For the VoIP we saw that the protocol uses UDP and the traffic, after a transition time, goes only between two hosts.
The average packet size is the MTU and the protocol tryies to exploit all the channel for a better QoE.
On another test, we tried the VoIP while a download was ongoing, and we saw that the quality
decreased to allow communication. The traffic is between the UDP port 80 \#\#TODO\#\# TOCHECK \#\#

The last traffic analysis we made was with a DOTA[explain what it is] web match.
We saw that for that type of gaming session, the packet size is about \# bytes and involves both TCP and UDP ports.
  \\
  \\
  For every networks performance we have tryed to develop a specific solution. The OpenFlow main goal is to centralize controll and software
  complexity so the routerboard can made only some easy function in orther to contain the hardware complexity and mantainly a good speed\cite{qos_paper}.
  Il che ha portato a dover utilizzare stratagemmi.
  \#TODO %riformulare
    \subsubsection{Delay}
Monitoring round-trip time provides important insights for network troubleshooting and traffic engineering.
The  common  monitoring  technique  is  to  actively  send  probe packets  from  selected  vantage  points  (hosts  or  middleboxes).
In traditional  networks, the control over the network routing is
limited,  making  it  impossible  to  monitor  every  selected  path. The  emerging  concept  of  Software  Defined  Networking  sim-plifies 
network  control.  However,  OpenFlow,  the  common  SDN protocol,   does   not   support   RTT   monitoring   as   part   of   its
specification. In this experience we have tryed to emply an tecnic to escape this problem\cite{rtt}.
\newline The idea is to measure the time that occure from sending a packet from controller to a switch and the time in which the packet return from 
an other switch.
\\
\newline First we have to discovery round trip time from controller to every single switch on our topology.
In order to do this on the controller we start a timer, we send a proper packet to the switch specifing the \texttt{OFPP_CONTROLLER - Send to the controller}
option and when the message return to the controller we stop the timer and calculare RTT for the switch.
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/rtt2.png}
 \caption{\emph{RTT between controller and OF switch}}
 \label{fig:topo}
\end{figure}
\newline
If we need to calculate the delay on a specific link, the rtt has to calculate with the headed switch of this link.
\\
\newline 
Now we create a proper message with unique identifier number, we start a timer \texttt{TM1} and then send this message to a first switch specifing the output port.
Then when the packet comes to the second switch it identify the particular message and sends to the controller that stops it's dedicated timer \texttt{TM2}.
\newline This part required a particolar flow entry on the second switch in orther to send to the controller only some particulare packet.
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/rtt1.png}
 \caption{\emph{Pck from controller to itself following a particular way.}}
 \label{fig:topo}
\end{figure}
\\
\newline At this point we can calculate the delay of the link.
\\
\\
 delay = (TM2 - TM1) - RTT1/2 - RTT2/2
\\
\\
Teorethically this method must works but in practice there are a lot of delay introduce from computational concestion on the virtualized controller.
What we observe is a estimated delay like some seconds that is impossible, infact if we made some \texttt{ping test} from one host to another host 
in this SDN network we realize delay around 3ms.
\\
